### 2023-05-12 analysis/paper/paper.qmd
Updated manuscript and restructured

---
title: "Quartzite Complexities: Non-destructive analysis of bifacial points from Västerbotten, Sweden"
author:
  - Mattias Sjölander:
      email: mattias.sjolander@umu.se
      institute: [UmU]
      correspondence: true
  - Johan Linderholm:
      email: johan.linderholm@umu.se
      institute: [UmU]
      correspondence: false
  - Paul Geladi:
      email: paul.geladi@slu.se
      institute: [SLU]
      correspondence: false
  - Philip I. Buckland:
      email: philip.buckland@umu.se
      institute: [UmU]
      correspondence: false
institute:
  - UmU: Department of Historical, Philosophical and Religious studies, Umeå University, Umeå, Sweden
  - SLU: Department of Forest Biomaterials and Technology, Swedish University of Agricultural Sciences, Umeå, Sweden 
title-block-published: "Last updated"  
date: now
date-format: long
crossref: 
  fig-title: '**Fig.**'
  fig-labels: '**arabic**'
  fig-prefix: Fig.
format:
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 300
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/quaternary-science-reviews.csl" # Insert path for the bib-style
abstract: |
  Northern Fennoscandia represents a geologically complex region affected by both glacial and postglacial processes. Provenance of locally available geological material thus becomes complex and fraught with uncertainty. Quartzite was a key material type utilized by hunter-gatherers in Northern Sweden around the 4 000 – 2 000 BP period, and is thus critical to the understanding of raw material procurement and material flow within the region. However, there is a severe lack of methodological development in the characterization of these materials. 125 points and preforms made from quartz/quartzite were sampled from 46 archaeological sites along the upper Ångerman river valley in Västerbotten, Sweden. The material has been analysed non-destructively using three separate portable spectroscopic instrumentations (Near-infrared, Raman, X-Ray Fluorescence). Evaluation of the spectra and exploratory data analysis using Principal Component Analysis demonstrates detectable differences in the material that likely stem from diagenetic/paragenetic origin. The presence of graphite, muscovite and biotite could likewise provide information on the material’s metamorphic grade. In addition to reaffirming the potential of field-based screening instrumentation, these results will benefit future surveys of geological sources in the region. They also indicate potential for the construction of a predictive model that could classify the quartzite based on its chemical characteristic. Such a model would prove useful in future spatial analysis and testing of models of raw material management.
keywords: |
  Lithics; Provenance; Quartzite; Spectroscopy; Sweden
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights:

* Non-destructive analysis of prehistoric bifacial points made in quartzite using near-infrared spectroscopy, Raman scattering and X-Ray Fluourescence. 
* Each method detects differences in the structural and chemical characteristics in the material, but on their own the resolution is too low for predictive modelling. 
* Data fusion techniques are recommended for predictive modelling using non-destructive spectroscopy on quartzite material.

<!-- The following code chunk defines some general settings how code chunks should behave. -->
```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300,
  encoding = "UTF-8"
)
```

<!-- The following function can be used to suppress message outputs -->
```{r}	
quietly <- function(x) {	
  sink(tempfile())	
  on.exit(sink())	
  invisible(force(suppressMessages(x)))	
}	
```

<!-- The following code chunk is used for inline calculations of samples and sites -->
```{r sample-count, results='hide'}
suppressPackageStartupMessages(library(tidyverse))

#Import descriptive metadata
metadata.csv <-
  read.csv2(here::here("analysis", "data", "raw_data", "metadata.csv"), sep = ";", dec = ".", colClasses=c("sample_id"="character"), header = TRUE, na = c("", "NA", "NULL"), encoding = "UTF-8")

#Filter the sample data to exclude reference samples
tot_samples <- 
  metadata.csv %>% 
  dplyr::filter( type == "Point" | type == "Point fragment" | type == "Preform", 
         material == "Brecciated quartz" | material == "Quartz" | material == "Quartzite")

#Create a vector containing all unique sample ID
tot_uniq_samples <- 
  unique(tot_samples$sample_id)

#Create a vector containing all unique site ID
tot_uniq_sites <-
  unique(tot_samples$site_id)

#Create a vector containing all unique sample ID made in breacciated quartz
bquartz <- unique(tot_samples %>% 
  filter(material == "Brecciated quartz"))$sample_id

#Create a vector containing all unique sample ID made in quartz
quartz <- unique(tot_samples %>% 
  filter(material == "Quartz"))$sample_id

#Create a vector containing all unique sample ID made in quartzite
quartzite <- unique(tot_samples %>% 
  filter(material == "Quartzite"))$sample_id
```

# Introduction

Petrographic characterization and provenance studies are instrumental for inferring trade and exchange among prehistoric societies [@tykot03]. The presence of lithic material foreign to the local region of an archaeological site, raises important questions on the processes that resulted in its eventual deposition and discovery. The characterization of geological material and identification of geological sources are the first steps in reconstructing a prehistoric community’s strategies for raw material selection and management. Flint and obsidian have been central to studies of trade and exchange in European prehistory, due in part to the well-surveyed geological sources for these materials [@tykot17]. This focus has led to a bias, however, which has been highlighted by Prieto et al. [-@prieto_etal19: 15]. Recent decades, however, have seen an increase in interest for non-flint lithics and raw material studies [@blomme_etal12; @dalpra_pitblado16; @prieto_etal19; @prieto_etal20; @prieto_etal21; @ramacciotti_etal19].

In Northern Sweden, quartz and quartzite dominate the lithic assemblages and have been used in settlement, exchange and mobility studies [@baudou78]. Models have been proposed for raw material sourcing and trade and exchange for the less common slate and flint finds within Northern Sweden [@halen94: 115-122; @lundberg97: 161-171]. There has, however, been little development in the petrographic characterization and provenance study of quartz and quartzite [@broadbent79; @callahan_etal92; @knutsson_etal16; @tallavaara_etal10]. ). A possible reason for this might well be the complex geological background of the region [@bargel03; @stephens_bergmanweihed20; @stroeven_etal16]. Northern Sweden is extensively covered by glacial till, and whilst there are identifiable source areas (e.g. small bedrock quarries), there is an extremely large number of other potential sources of raw materials to be found in moraine deposits. The few existing theories on lithic material sourcing within the North Swedish region describe strategies for direct procurement from the mountain zone [@forsberg85; -@forsberg12]. This does not, however, account for the possibility that the material was sourced locally around the site, as part of an embedded adaptation strategy [@binford79; @rankama_etal06: 249-250]. A new approach to provenance studies is therefore required in this region.

Prieto et al. [-@prieto_etal19: 15; -@prieto_etal20: 32] note the ambiguous usage of the term quartzite in archaeology, with it often being used interchangeably between sedimentary orthoquartzite and metamorphic quartzite. When defining quartzite, different researchers have emphasised different features as being characteristics of the material [@howard05]. Visual characterization with the naked eye is typically unreliable, and microscopic analysis is recommended for the identification of metamorphic structures. The need for field-based methodologies is, however, stressed by Howard [-@howard05: 708]. Spectroscopic surface measurements are fast and can be adapted for use in the field. However, artefact collections (museum, private, etc.) also represent a valuable resource and asset, with easy access to large potential datasets with a wide geographical distribution. Destructive analysis on material from these collections, which are often unique and include sensitive cultural heritage material, are not always suitable or permitted, and thus there is a clear need for the development and testing of non-destructive techniques such as those presented in this study.

In this study, three different spectroscopic techniques have been applied to material stored in collections at Västerbottens museum in Umeå, Sweden. X-Ray Fluorescence (XRF) is a well-established spectroscopic technique for petrographic characterization [@shackley11a; @tykot03], due to its capability of providing element compositional data. The use of near-infrared spectroscopy (NIRS) is similarly well-known, as described by Norris [-@norris96], and there is a large body of work dedicated to the fingerprinting of geological materials [@hunt77; @clark_etal90; @linderholm_geladi14a; @linderholm_geladi14b; @sciuto18]. Complementary to NIR, Raman spectroscopy has been applied extensively in both the Cultural Heritage sector, in the analysis of artwork and manuscripts, as well as the Earth sciences [@dubessy_etal12].

# Aims {-}

This paper presents an assessment of the potential for non-destructive characterization of quartz and quartzite material without the use of references from known geological sources. The geological complexity of both the region and the material necessitates alternative approaches to studies of raw material management. Without relying on well-delimited source areas for geological material, a predictive model capable of separating different “classes” of quartzite would enable spatial analysis of settlement and mobility, as well as testing of established models. The currently available data on the chemical characteristic of quartz and quartzite in this region of Northern Sweden is limited, however, and this study represents a first step towards developing the topic of raw material management of these materials in Northern Sweden between 4 000 and 2 000 BP.

# Material {-}

\linebreak
 \linebreak

This study includes spectroscopic surface measurements collected on bifacial points (probable arrowheads and spear points) made of quartz and quartzite. The points have been recovered from archaeological sites during excavations and surveys, including stray finds, in Västerbotten County, Sweden (@fig-distribution-map). The sites are distributed along the Ångerman river valley, from Lake Kultsjön in the north, towards the town of Åsele in the south. The dataset consists of `r length(tot_uniq_samples)` points, point fragments and preforms stored in the collections at Västerbottens museum (www.vbm.se) and have been analysed using NIRS, Raman scattering and XRF.

The geographically closest chronology for bifacial points is centred around southeast Norway [@mjaerum12], but a few attempts have been made in the North Swedish region [@baudou92: 99; @forsberg85: 5; @forsberg89]. The latter suggest a period of use between ca. 3 800 - 2 000 BP [@kaeck09: 60]. Studies of raw material use at sites dating to this period seem to indicate an increase in the use of quartzite when compared to earlier periods [@baudou78].

\linebreak

```{r}
#| label: fig-distribution-map
#| cache: false
#| cache-lazy: false
#| fig-cap: "Geographical distribution of the sampled sites along the river of Ångermanälven, Västerbotten, Sweden. The glacial disturbance is visible in the widely distributed moraine. Tectonic symbology after [@gee_stephens20]. Made with © Sveriges Geologiska Undersökning, Lantmäteriet (CC0), Natural Earth, and Riksantikvarieämbetet"

knitr::include_graphics(here::here("analysis", "figures", "site_distribution_v1.jpg"), error = FALSE)
```

\linebreak

The dataset includes three main groups of petrographic material, as classified by archaeologists: quartz, quartzite, and brecciated quartz (Swe. "*brecciekvarts*"). Artefacts that had no prior classification were sorted into either the quartz or quartzite category based on visual inspection. This resulted in: `r length(bquartz)` brecciated quartz samples, `r length(quartz)` quartz samples, and `r length(quartzite)` quartzite samples. The inclusion of quartz and brecciated quartz in the study enables the separation of quartz from quartzite, based on non-destructive analysis of their chemical and structural characteristics, to be evaluated.

In his analysis of the petrographic material in the Early Norrland project [@biornstad68; @ahman67], the geologist Åhman [-@ahman67: 8] introduced the term *brecciated quartz* (also referred to as *mylonite quartz*) to archaeologists in Northern Sweden. Åhman describes the material as easily mistaken for an unusually pure quartzite, but where the fragmented configuration of the quartz can be identified via microscopic analysis. *Brecciated quartz* in particular became a common term used in North Swedish archaeology, and has typically been used to identify a certain translucent type of fine-grained, quartz-rich material [e.g. @holm91: 24]. As the material frequently has been characterized on the basis of macroscopic characteristics, despite the fundamental features necessary for characterization being microscopic, there is value in including this material as a test group in the study.

\linebreak

```{r}
#| label: fig-artefacts
#| cache: false
#| cache-lazy: false
#| fig-cap: "Examples of the bifacial points studied. The material is heterogeneous in terms of colour, and the surface of the artefacts themselves is not uniform. Ranging from near perfect points to rough preforms, it can be difficult to sample due to light scattering effects and saturation. Length varies between 12 – 138 mm, width between 8 – 67 mm, and thickness between 3 – 76 mm"

knitr::include_graphics(here::here("analysis", "figures", "figure_2.jpg"), error = FALSE)
```

\linebreak

# Methods {-}

\linebreak
 \linebreak

All measurements were collected at the Environmental Archaeology Lab at Umeå University (www.umu.se/en/research/infrastructure/mal/). The analysis was conducted in a dark room in order to minimize stray light, with the main light sources in the room being 2 computer screens. The sampled material is heterogeneous in colour and shape (@fig-artefacts - @fig-artefact-size) (see supplementary data). The artefacts range from small and flat point fragments, to large and rough preforms. Therefore, the signal integration could at times be difficult, leading to scattering effects and saturation. For NIR and Raman analyses, each artefact was sampled twice, on different sides, with additional measurements where sampling was difficult. When sampling the probe was in contact with the surface of the artefact at as flat a surface as possible. If the material featured high heterogeneity, the sample point was focused on the surrounding matrix, avoiding deviating inclusions (e.g. the dark mineral veins in @fig-artefact-size).

For XRF, the more time-consuming nature of the measurement process necessitated only one measurement per object. The representativeness of the XRF analysis is thus lower, and inferences derived from the results in this study should be used with caution. The risk of misinterpretation is, however, somewhat compensated by the large size of the dataset, with larger sites including multiple objects that may come from the same geological source. The results from sites with smaller assemblages are, however, more liable to be affected by the limited range of material analysed.

In colour, the artefacts exhibit a mainly black or grey hue, with a subset of semi-translucent material with a beige tint, with some being almost fully translucent. While the colour of the material has been classified according to the Munsell system (see supplementary data), for visualization purposes they have also been sorted into general hues: *dark* (~ N1-N4), *light* (~ N5-N8), *white* (~ N9), and *colourless* (lack of colour, opaque to translucent). In the case of heterogeneous objects the Munsell colour and general hue describes the surrounding matrix.

\linebreak

```{r}
#| label: fig-artefact-size
#| cache: true
#| fig-width: 7.87
#| fig-asp: 0.618
#| fig-align: center
#| fig-cap: "Scatter plot of length and width measurements (mm) for the sampled artefacts. 8 artefacts have no recorded thickness, which ranges between 3-76 mm. Note that the term point does not differentiate between spear points or arrowheads"
quietly(source(here::here("analysis", "paper", "001-artefact-size.R"), encoding = "UTF-8"))
#knitr::include_graphics(here::here("analysis", "figures", "001-artefact-size.jpeg"), error = FALSE)

fig
```

\linebreak

## Instrumentation and data pre-processing {-}

\linebreak
 \linebreak

### Near Infrared Spectroscopy (NIRS) {-}

Many chemical bonds between small atoms (H, C, O, N etc.) show a fundamental vibration in the near-infrared region (780 nm – 2 500 nm) of the electromagnetic spectrum. The vibration occurs at a frequency that is typical for the bond and can be used to identify and quantify it. Based on the theory of molecular vibration and excitation [@hunt77], by exposing a geological material to electromagnetic radiation it is possible to generate overtone vibrations which can be related to certain molecular groups. The most common group capable of generating these overtones is the OH group, which in quartz-based materials typically relates to molecular water and metal-OH combinations [@hunt77: 508]. Overtones occur at frequencies that are a multiple of the infrared vibration, and most of these end up in the near-infrared region [@osborne_etal93].

NIR analysis was conducted with an Analytical Spectral Device (ASD) LabSpec 4, using a contact probe (spot size: 10 mm). The ASD features a detection range of 350 nm – 2 500 nm, with spectral sampling (resp. spectral resolution) of 1.4 nm (resp. 3 nm) in the visible and near infrared range and 1.1 nm (resp. 10 nm) in the short‐wave infrared range. Each object was measured against a white reference background and the collected measurements were averaged for each artefact. The data was restricted to the 1 000 nm - 2 500 nm range and mean-centred before Principal Component Analysis (PCA) modelling.

\linebreak

### Raman {-}
Raman spectroscopy shares similarities with NIRS in that they both can be used to infer molecular structure based on their vibrational properties within the near-infrared region [@nafie01]. Raman spectroscopy is based on the *Raman scattering* effect. By exposing a molecule to radiation with a monochromatic laser it is possible to induce a dipole moment in the molecule which generates the Raman effect. The subsequent light that is scattered contains both Rayleigh and Raman scatter, the former at a frequency of incident radiation and the latter at a shifted frequency [@dubessy_etal12]. The spectrometer will measure the Raman scatter and generate a spectrum where the spectral line has shifted compared to the incident Rayleigh frequency. As the structure of a molecule will inform its vibrational mode, the point at which these shifts occur within the infrared region varies depending on the material [@smith_carabatos-nedelec01].

The Raman instrumentation used was a portable i-Raman EX featuring a 1064 nm excitation laser. The spectral coverage of the instrumentation is 175 cm^-1^ – 2 500 cm^-1^, with a resolution of 9.5 cm^-1^. Integration time was set to 20 seconds with 30% power. After averaging the spectra for each artefact the results were baseline corrected in order to detrend the data. This was achieved using Lieber and Mahadevan-Jansens' [-@lieber_mahadevan-jansen03] algorithm for polynomial fitting, with the ChemoSpec package [@hanson16]. The spectra were then mean-centred before PCA.

\linebreak

### X-Ray Fluorescence (XRF) {-}
XRF spectroscopy is a well-established method in petrographic characterization in archaeology [@hall60; @shackley11a]. Spectral features in XRF spectrometry can be related to certain elements according to Moseley’s law [@moseley1913], which recognizes that the energy of the fluorescence produced when a material is irradiated with high energy particles is characteristic to certain atoms [@thyrel14: 20-24]. Quantification based on these features is complex, however, and a number of methods have been developed as a result [@sitko_zawisa12: 143-159]. In a non-destructive context, where the sample is not homogeneous, it is necessary to correct for matrix effects caused by the presence of other elements in a sample. One way to do so is calibration using pure-element standards, with which a calibration constant can be calculated.
XRF measurements were collected using a Thermo Scientific Niton XL5 Analyzer (spot size: 8 mm), connected to a portable test stand. The reference calibration used for element quantification was mining mode [@knight_etal21: 2]. The integration time was set to 130 seconds.

As the variables in XRF analysis are given as proportions of a constant sum (e.g. percentage, parts per million (ppm)), this creates something referred to as the closed sum effect [@pawlowsky-glahn_egozcue06]. This means that if one element was to change in value, others would follow to maintain the constant [@grunsky_caritat20]. This has implications for classical statistical methods and is managed by transforming the data using log-ratio transformation [@aitchison99]. Currently there are three main methods; additive log-ratio (alr), centered log-ratio (clr) and isometric log-ratio [@pawlowsky-glahn_etal15]. In the PCA, the Principal Components are orthogonal and thus the relationships between variables treated with log-ratio transformation are linear and independent of the total concentration [e.g. @reimann_etal12]. With this transformation, PCA is thus suitable for exploring geochemical trends related to the rock-forming process without being impacted by the closed sum effect (REF).

As the current dataset features values below the Limit of Detection (LOD) these were treated as NULL (or “missing values”) during PCA. A LOD value is defined as a measured value 3 times lower than the standard deviation for that measurement. Missing values are not permitted in the PCA and thus had to be imputed before the analysis, which was performed using the *lrEM* function from the *zCompositions* R package [@palarea-albaladejo_martin-fernandez15]. *lrEM* stands for “Log-ratio Expectation-Maximisation”, which is an algorithm used to impute “left-censored data” (e.g. values below LOD, rounded zeros). Following log-transformation the data was mean-centred and scaled to unit variance before PCA. Samples that were less than 10 mm in length or width were excluded from the model [@shackley11b: 9].

\linebreak

## Principal Component Analysis (PCA) {-}

The data were evaluated using PCA, an unsupervised exploratory technique for reducing the dimensionality of a dataset  [@geladi_linderholm20]. Multivariate approaches, such as PCA, have been proven useful for exploring complex spectroscopic datasets [@sciuto18]. Clusters and trends that emerge in the resulting score plot can potentially be related to the spectral fingerprints of the material. These groupings can then be followed up with supervised methods, and aid in the construction of predictive models for classifying unknown materials according to previously defined groups.

PCA analysis was performed using the base R *prcomp()* function [@Rteam21a; @Rteam21b].

\pagebreak

# Results {-}

Spectra and PCA loadings for each instrumentation can be found in the supplementary data. An open repository with the code used in the analysis is linked in the *Data availability* section.

\linebreak

## Near Infrared Spectroscopy (NIRS) {-}
```{r nir-sample-count, results='hide'}
#Select all dark samples
points.d.tot <-
  metadata.csv %>%
  filter( type == "Point" | type == "Point fragment" | type == "Preform", 
         material == "Brecciated quartz" | material == "Quartz" | material == "Quartzite") %>% 
  filter(hue == "Dark")

#Filter out the dark quartzite samples that feature peaks of interest in their spectra 
points.d.filt <-
  points.d.tot %>% 
  filter(!sample_id %in% c("153","167","168","169","172","174","175","176","177","182","190","191","193","194","196","198","200","204","214",
                           "215","216","229","238","262","265","272","282","359","391","392","393","397","405",
                           "408","410","411","413","414","415","416","417","424","425","428","430","432","55","56"))

#pass to vector so that length() can be used in text to quote total nr. of samples
spec.d.total <- 
  points.d.tot$sample_id

#pass to vector so that length() can be used in text to quote nr. of dark samples after SG analysis
spec.d.filtered <- 
  points.d.filt$sample_id
```

NIRS is less effective on translucent material, such as a pure quartz crystal, as the absence of impurities in the material will result in little to no reflection/absorbance of light [@hunt77]. Similarly, it can be difficult to work with dark materials as most of the light produced by the instrument will be absorbed, with the high absorbance potentially obscuring visible features in the spectra. The dark material exhibits a wider range of absorbance, most likely due to the range of hues within the group, and generally shows less distinct peaks than the material with a lighter hue.

The spectra were evaluated using Savitzky-Golay (SG) filtering, a smoothing technique widely applied in spectroscopy [@chryssikos_gates17: 78-81]. 2^nd^ derivative SG can amplify weaker peaks in the spectra, although at the risk of amplifying noise [@mark_workman21: 351-397; @schmid_etal22]. After reviewing the *dark* group using SG filtering, `r length(spec.d.filtered)` out of a total of `r length(spec.d.total)` artefacts were kept in the analysis.

Significant peaks were identified at 1 400 nm, 1 900 nm, 2 200 nm, and 2 350 nm. These features are all the result of overtones and vibration combinations of hydroxyl (OH) groups, with the two first peaks at 1 400 nm and 1 900 nm being indicative of molecular water (@fig-nir-peak-summary) [@hunt77: 508-510; @sciuto_etal19]. OH groups can produce a response in a number of different fields in the same material, with metal-OH combinations occurring in the 2 000 nm - 2 500 nm region. While there are exceptions, a more intense 2 200 nm band has typically been related to AlOH, and a more intense 2 350 nm band to MgOH [@hunt77: 509; @clark_etal90: 12664]. In the current dataset the 2 200 nm band is typically the more intense band.

The 2 200 nm and 2 350 nm absorption bands can be found in mica, such as muscovite and biotite, which can be found in quartzite [@clark_etal90: 12; @hunt77: 510; @ramanaidou_etal15: 207]. Although muscovite and biotite share some band positions, the altered Al content as a response to metamorphism has been documented to shift some positions and increase others, such as the 2 250 nm band for biotite [@duke94]. As can be seen in the average spectra in @fig-nir-spectra-summary, the dark material exhibits an additional peak at ~2 250 nm, which may indicate the presence of biotite and a different metamorphic grade.

Two smaller features were recorded, albeit vague and difficult to make out in the non-processed spectra, at 1 065 nm and 1 300 nm (@fig-nir-spectra-summary). These could be caused by ferrous iron (Fe^2+^) [@hunt77: 503-504], which has been demonstrated to produce a response in the NIR in a number of minerals. In studies of olivine the absorption region has been determined to be broad and centred around 1 000 nm, but with three features at 900 nm, 1 100 nm and 1 300 nm [@trang_etal13]. Alternatively, weak water features have also been noted in the 950 nm and 1 150 nm bands, particularly among clays with adsorped water. However, these overlap with the Fe^2+^ bands in, for example, biotite [@adams75: 104-106]. It is possible that the features at 1 065 nm and 1 300 nm can be linked to the presence of ferrous iron, but it is difficult to say what the contributing mineral is at this stage.

\linebreak
<!--
```{r fig.width=12, fig.height = 5}
#| label: fig-nir-spectra-summary
#| cache: true
#| fig-width: 7.87
#| fig-asp: 0.8
#| out-width: "70%"
#| fig-align: center
#| fig-cap: "Summary of the raw spectra for each colour group. Note the differences in absorbance between the groups, with the dark group featuring a larger range of absorbance values as well as less distinct peaks."

quietly(source(here::here("analysis", "paper", "001-nir-spectra-summary.R"), encoding = "UTF-8"))
#knitr::include_graphics(here::here("analysis", "figures", "001-nir-spectra-summary.jpeg"), error = FALSE)

fig
```

\linebreak

```{r fig.width=12, fig.height = 5}
#| label: fig-nir-savitzky-transform
#| cache: true
#| fig-width: 7.87
#| fig-asp: 0.8
#| out-width: "70%"
#| fig-align: center
#| fig-cap: "Comparison of NIR spectra between two samples from the dark group. a) the non-processed spectra for sample 258 and 411, b) sample 258 spectra with 2^nd^ derivative Savitzky-Golay filtering, c) sample 411 spectra with 2^nd^ derivative Savitzky-Golay filtering."

quietly(source(here::here("analysis", "paper", "003-nir-savitzky-transform.R"), encoding = "UTF-8"))
#knitr::include_graphics(here::here("analysis", "figures", "003-nir-savitzky-transform.jpeg"), error = FALSE)

fig
```

\linebreak

```{r}
#| label: fig-nir-savitzky-transform-alt
#| cache: true
#| fig-width: 7.87
#| fig-asp: 0.6
#| out-width: "70%"
#| fig-align: center
#| fig-cap: "Savitzky-Golay transformed spectra contrasting the results between the spectra in the dark (black) and light (orange) groups."

quietly(source(here::here("analysis", "paper", "003-nir-savitzky-transform-alt.R"), encoding = "UTF-8"))
#knitr::include_graphics(here::here("analysis", "figures", "003-nir-savitzky-transform-alt.jpeg"), error = FALSE)

fig
```
-->

```{r}
#| label: fig-nir-peak-summary
#| cache: true
#| fig-width: 7.87
#| fig-asp: 1
#| out-width: "70%"
#| fig-align: center
#| fig-cap: "The main features in the NIR spectra can be related to overtones and vibration combinations of OH groups. The 1 400 nm and 1 900 nm peaks are indicative of molecular water, and common in quartz-based materials. The 2 200 nm and 2 350 nm bands indicate metal-OH combinations, where a more intense 2 200 nm band has typically been related to the presence of AlOH, and a more intense 2 350 nm band MgOH. Although difficult to detect in the non-processed spectra, weaker features also appear at 1 065 nm and 1 300 nm, that may be the result of ferrous iron (Fe^2+^)"

quietly(source(here::here("analysis", "paper", "002-nir-peak-summary.R"), encoding = "UTF-8"))
#knitr::include_graphics(here::here("analysis", "figures", "002-nir-peak-summary.jpeg"), error = FALSE)

fig
```

\linebreak

```{r}
#| label: pca-var
#| cache: true

quietly(source(here::here("analysis", "paper","004-nir-pca.R"), encoding = "UTF-8"))
nir.pc1var <- nir.pc1var
nir.pc2var <- nir.pc2var
nir.pc3var <- nir.pc3var
```

\linebreak

The loading lines for the PCA can be seen in the supplementary data. The first principal component (PC 1) shows a strong trend (`r nir.pc1var` %) that seems to largely correlate with the absorbance of the material (@fig-nir-pca). This can be seen in the trend of darker material clustering towards the positive side of PC 1, with lighter towards the negative. A slight difference in the clustering of material with a light hue can also be seen along PC 2 (`r nir.pc2var` %). Objects with a white or light grey hue gravitate towards the bottom left of the score plot, while material that displays some translucence cluster towards the top left. There are objects that break this trend, however, such as sample 255 in the bottom right of @fig-nir-pca :A.

Whilst the explained variance of PC 3 is low, (`r nir.pc3var` %), it still contains a number of the aforementioned features visible in the spectra (inverse to PC 1). Adding PC 3 to a score plot (@fig-nir-pca :B), however, provides little new information apart from a couple of samples, mainly from the light group, that break away from the larger cluster.

\linebreak

```{r}
#| label: fig-nir-pca
#| cache: true
#| fig-width: 7.87
#| fig-asp: 1
#| out-width: "70%"
#| fig-align: center
#| fig-cap: "NIR PCA score plot for PC 1-3: a) dark and light material is separated out along PC 1, due to absorbance, with white and colourless material separating along PC 2, b) PC 3 adds little new information, mainly separating three of the light samples"
#| 
quietly(source(here::here("analysis", "paper","004-nir-pca.R"), encoding = "UTF-8"))
#knitr::include_graphics(here::here("analysis", "figures", "004-nir-pca-with-photos.jpg"), error = FALSE)

fig
```

\linebreak

## Raman {-}
There are two distinguishing features present in the Raman data (@fig-raman-peak-ident)), one which can be related to quartz and another that may indicate the presence of carbon (C) structures. The three initial bands at ~ 128 cm^-1^, ~ 200 cm^-1^ and ~ 460 cm^-1^, , are some of the more intense quartz bands [@zhong_etal21], with the strongest peak occurring at 460 cm^-1^ [@gillet_etal90].

The second feature shows as two peaks at around 1 287 cm^-1^ and 1 598 cm^-1^, and dominates the spectra in terms of intensity for the darker samples. These positions fall within what is referred to as the *D* and *G* band respectively, and have been observed in spectroscopic studies of graphite [@cesare_maineri99]. The D band occurs as a result of disordered carbon (C) structures, whilst the G band is indicative of sp^2^ carbon networks and caused by stretching in the carbon bonds, which is common in graphite-based materials [@cohen-ofri_etal06; @vollebregt_etal12; @lee_etal21]. The G band peak falls within the expected position, whereas the position of the D band is more commonly identified as being in the 1 330 - 1 350 cm^-1^ region. It has, however, been documented that the dispersive behaviour of the D band can cause a frequency change as a result of the energy of the incident laser [@pimenta_etal07: 1280]. This could thus explain the discrepancy between the observed peak in the dataset and the literature.

In addition to the more intense features described above, a number of weaker peaks can also be seen in the spectra. Some of these could be additional quartz bands [@zhong_etal21], such as a weak peak at 1160 cm^-1^ that is somewhat obscured by the D band. that is obscured by the D band. Two peaks can also be found in the 700 - 800 cm^-1^ band range, and while the 700 cm^-1^ band has been identified as one of the aforementioned quartz bands, studies of clay minerals [@kloprogge17], and iron ore minerals [@ramanaidou_etal15], have also generated a response related to Si, Al and OH vibrations within the 700 - 800 cm^-1^ region.

\linebreak
<!--
```{r}
#| label: fig-raman-spectra-summary
#| cache: true
#| fig-width: 7.87
#| fig-asp: 0.8
#| out-width: "70%"
#| fig-align: center
#| fig-cap: "The non-processed Raman spectra visualised by colour group. Note that the dark material in general features more intense peaks at wavenumber ~ 1 300 - 1 600 (cm^-1^)"

quietly(source(here::here("analysis", "paper", "005-raman-spectra-summary.R"), encoding = "UTF-8"))
#knitr::include_graphics(here::here("analysis", "figures", "005-raman-spectra-summary.jpeg"), error = FALSE)

fig
```
-->
 \linebreak

```{r}
#| label: raman-baseline
#| cache: true

#baseline correct raman data
quietly(source(here::here("analysis", "paper", "006-raman-baseline.R"), encoding = "UTF-8"))
```

```{r}
#| label: fig-raman-peak-ident
#| cache: true
#| fig-width: 7.87
#| fig-asp: 0.8
#| out-width: "70%"
#| fig-align: center
#| fig-cap: "TAveraged and baseline corrected Raman spectra for the light and dark groups. The features can be related to quartz and carbon structures in the material. The initial three peaks in 100 - 500 cm^-1^ range correspond to some of the more well-known vibration modes of quartz. The 1 287 cm^-1^ and 1 598 cm^-1^ peaks fall within what is referred to as the D and G band respectively, indicating the presence of carbon structures"

#generate figure
quietly(source(here::here("analysis", "paper", "006-raman-peak-identification.R"), encoding = "UTF-8"))
#knitr::include_graphics(here::here("analysis", "figures", "006-raman-peak-identification.jpeg"), error = FALSE)

fig
```

\linebreak

```{r}
#| label: raman-var
#| cache: true

quietly(eval(parse(here::here("analysis", "paper", "007-raman-pca.R"), encoding = "UTF-8")))
raman.pc1var <- raman.pc1var
raman.pc2var <- raman.pc2var
raman.pc3var <- raman.pc3var
```

The PCA model (@fig-raman-pca) shows a clear trend along PC 1 (`r raman.pc1var` %) related to the quartz and D/G peaks. Samples where the D/G peaks dominate in terms of intensity cluster towards the positive side of PC 1, with those that completely lack or have weaker D/G bands on the opposite side. Similarly, along PC 2 (`r raman.pc2var` %) the samples that have higher intensity D/G peaks gravitate towards the top of the score plot. Worth noting are two dark samples that completely lack D/G peaks, placing them among the light material in the left side of the plot. One of these is a brecciated quartz sample with a dark yellow tint, and thus more a sign of the difficulties with assigning a generic hue to the material. The second sample, however, stands out from the rest of the material. It is a point preform made from a grey material with a rough texture, and possible layering, with larger quartz crystals. It does not fit with the “expected” image of a quartzite and may be a case of erroneous classification. The addition of PC 3 (`r raman.pc3var` %) provides no new information, but rather separates out those samples with a higher level of noise.

\linebreak

```{r}
#| label: fig-raman-pca
#| cache: false
#| cache-lazy: false
#| fig-cap: "PCA score plot for Raman data. Objects that feature only quartz peaks are clustered towards the top left of the plot, whilst the darker material that feature dominant D/G peaks are distributed towards the right of the plot. In the middle cluster the samples feature both quartz and D/G peaks, with those that have higher intensity D/G peaks in the bottom of the cluster"

#quietly(source(here::here("analysis", "paper", "007-raman-pca.R"), encoding = "UTF-8"))
knitr::include_graphics(here::here("analysis", "figures", "007-raman-pca-with-photos.jpg"), error = FALSE)
```

\linebreak

## X-Ray Fluorescence (XRF) {-}

```{r}
quietly(source(here::here("analysis", "paper", "008-xrf-violin-plot.R"), encoding = "UTF-8"))
#| label: xrf-si
#| cache: true

#extract samples above 80% Si 
Si_90 <- Points.xrf %>% 
  filter(`Si` > 90)

#calculate percentage of samples above 80% Si
si_90 <- length(unique(Si_90$sample_id))
```

```{r}
#| label: xrf-pca-var
#| cache: true

quietly(source(here::here("analysis", "paper", "009-xrf-pca.R"), encoding = "UTF-8"))

xrf_var <- xrf.pc1var + xrf.pc2var

xrf_var_low <- xrf.pc1var.low + xrf.pc2var.low + xrf.pc3var.low + xrf.pc4var.low
```

The XRF spectrometer is not capable of detecting all the common elements, and the software therefore calculates a balance (BAL) that represents the remaining unquantifiable elements in a sample. The elements represented in this group are all those with an atomic number lower than ~12~Mg [@thermo18]. The analysis shows that ~14~Si makes up the main main bulk of the material, with ~13~Al recorded up to ~ 5% for some samples (@fig-xrf-violin-plot). The majority of the readings resulted in ~14~Si content of less than 50%, as well as a BAL between 40 – 70%, meaning that a significant portion of the material remains unknown. Whilst this could be due to the limitations of the instrumentation, it could also be a result of the sampling. As quartzite is a metamorphic rock where the distribution of quartz is not necessarily uniform in the material, the sample point can have a considerable influence on the measurement. Out of a total of `r length(tot_uniq_samples)` samples `r si_90` measured a ~14~Si content higher than 90%. ~12~Mg, ~23~V, ~25~Mn, ~39~Y and ~30~Zn recorded a high number of values below the LOD ("missing values"), although in the case of ~12~Mg and ~25~Mn the recorded values reached up to 0.3% and 3% respectively.

Initial analysis used clr transformed data, but encountered issues related to the structure of the ~14~Si values. Instead log10-transformation was used, producing a similar model, but one that managed the loading for ~14~Si better. The PCA was run twice, firstly with the > 90% ~14~Si samples included and secondly excluded. The initial PCA (model A), with the full dataset included, resulted in 3 PCs with an eigenvalue above 1, and a cumulative explained variance of `r xrf_var`%. K-means cluster analysis was performed on the PCA data in order identify potential groups with similar elemental composition. Using the elbow graph method 3 clusters were deemed suitable for the analysis.

\linebreak

```{r}
#| label: fig-xrf-violin-plot
#| cache: true
#| fig-width: 7.87
#| fig-asp: 1
#| out-width: "70%"
#| fig-align: center
#| fig-cap: "Violin plots of the XRF values per element. Median and interquartile range is represented with a boxplot, with the half-violin showing kernel density estimation. Number of values above LOD is noted by n. Of particular note is the bimodal distribution of Si values. Y and Zn were left out as they had less than 30 values above LOD"

quietly(source(here::here("analysis", "paper", "008-xrf-violin-plot.R"), encoding = "UTF-8"))

fig
```

\linebreak

As can be seen in the score plot (@fig-xrf-pca :A), the three K-means clusters remain separated. Cluster 1 is made up of those samples that feature a high ~14~Si content (> 90%). There is no real pattern to this cluster aside from higher ~14~Si values with lower ~26~Fe and ~40~Zr values. Loadings for elements expected to follow changes in 14Si content also end up near cluster 1 (e.g. Al, Ca). One of the “quartz” labelled samples ends up in cluster 3 rather than 1, and upon examining the sample visually it does not display any typical quartz characteristics.

Cluster 2 is distinguished by a higher ~22~Ti, ~26~Fe and ~40~Zr content. It mainly features samples with a dark hue, with the few samples from the light group being grey in colour and arguably somewhere in between a “dark” or “light” hue. ~13~Al and ~19~K is similarly higher in cluster 2 than cluster 3, which is in turn marked by its, on average, lower elemental content than the other two clusters. As ~13~Al and ~19~K are indicative of a sedimentary depositional context that has been subject to metamorphism (REF FROM JOHAN), and the cluster with higher values may thus indicate samples with a different grade of metamorphism. Cluster 3 features a more limited range of values between the samples and includes all the brecciated quartz.

In the second PCA model (model B) four PCs feature an eigenvalue above 1 with a cumulative explained variance of `r xrf_var_low`%. Three clusters were deemed suitable for the K-means analysis (@fig-xrf-kmeans-boxplot). The range of values is more limited in cluster 3, with overall lower values than compared to the other clusters (@fig-xrf-pca :B). The *brecciated quartz* samples all remain in cluster 3, along with mainly light-to-colourless samples. Cluster 2 features the main bulk of darker samples with higher  ~22~Ti, ~26~Fe and ~40~Zr values. Cluster 1 likewise feature a large portion of the darker materials, with comparable ~22~Ti, ~26~Fe and ~40~Zr values. It also has higher values of ~12~Mg, ~13~Al, ~16~S, ~19~K, and ~20~Ca.

~38~Sr was detected at trace amounts. It is a lithophile element that can substitute for 20Ca in sulphates, feldspars, mica and clays [@ahijado_etal05;@banner95], which might be one explanation for its presence in cluster 1 in model B. Nearly all samples with ~38~Sr are located in cluster 2 in model A.

```{r}
#| label: fig-xrf-pca
#| cache: true
#| fig-cap: "XRF PCA score plots: a) Score plot featuring the full dataset, cluster 1 includes all samples with Si > 90%, while the remaining samples seem to cluster based on Fe and Zr content. This results in most of the darker materials clustering towards the bottom of the plot. b) Score plot with > 90% Si samples excluded. Similar to plot A there is a separation based on Fe and Zr values, with darker material towards the bottom of the plot. Light-to-colourless samples remain largely together in cluster 3. Cluster 1 stands out by featuring slightly lower Si, Fe, Ti, and Zr values, but higher values in Al, Ca, K, and Mg"

knitr::include_graphics(here::here("analysis", "figures", "009-xrf-pca-with-photos.jpg"), error = FALSE)
```

\linebreak

```{r}
#| label: fig-xrf-kmeans-boxplot
#| cache: true
#| fig-width: 7.87
#| fig-asp: 1
#| out-width: "100%"
#| fig-cap: "Boxplot of XRF values (%) in PCA model B, grouped by K-means cluster. Cluster 1 is made up of samples with lower values of Si, but higher values for those elements that are typically associated with Si (e.g. Al, Ca). Cluster 2 includes the samples with higher Fe/Ti/Zr content, whilst cluster 3 is mainly distinguished by a higher Si content and low Fe/Ti/Zr content"

#Since knitr commands for adjusting knitr imported figures don't work in quarto, and the xrf script generates multiple figures, the easiest way to manage this is saving the boxplot
#as an object to derived_data when running xrf-pca-var above and then loading it here and outputting it in manuscript
fig <- 
  readRDS(file = here::here("analysis", "data", "derived_data", "009-kmeans-box-modelb.rds"))

fig
```

\linebreak

The K-means cluster analysis of model B is visualised in the NIRS PCA score plot, which further emphasizes the variation within the general colour groups (@fig-nir-xrf-kmeans). Cluster 2, with higher ~26~Fe and ~40~Zr values, mainly coincides with the darker material to the right side of the plot. This was expected as the trend along PC 1 in the NIR PCA mainly separates the material according to absorbance. Cluster 3 in turn coincides with the lighter material in the opposite side of the model. Cluster 1, however, is fairly evenly distributed within the model along PC 1 and PC 2.

The variation that occurs within the general colour groups can be further exemplified by looking closer at three samples selected from the dark group, that each belong to a different K-means cluster. They all share a similar hue with D/G peaks detected in the Raman analysis (@fig-nir-sample-comparison). All three samples have similar absorbance values, but sample 391 exhibits differences in the spectra when compared to the other two samples. Samples 386 and 404, however, are separated primarily in the XRF analysis.

```{r}
#| label: fig-nir-xrf-kmeans
#| cache: true
#| fig-width: 10
#| fig-asp: 0.687
#| out-width: "70%"
#| fig-cap: "NIR PCA score plot for PC 1-2, with the K-means cluster from XRF PCA model B visualised in the colour of the points. Red points are samples that were excluded in the XRF PCA. The trend separating dark-light samples in the NIR PCA mostly keeps cluster 2-3 separated, showing some overlap in the results, but cluster 1 is fairly evenly distributed in the score plot. This means that whatever differences can be discerned in the compositional data does not seem to be detectable using the ASD"

#Since knitr commands for adjusting knitr imported figures don't work in quarto, and the xrf script generates multiple figures, the easiest way to manage this is saving the boxplot
#as an object to derived_data when running xrf-pca-var above and then loading it here and outputting it in manuscript
fig <- 
  readRDS(file = here::here("analysis", "data", "derived_data", "009-kmeans-nir-pca.rds"))

fig
```

\linebreak

```{r}
#| cache: true
#Quarto has issues sizing the figure, perhaps related to the fact that it is made by turning multiple plots into rasters. Therefore the base figure import is used here, and not in any other part of the document.
quietly(source(here::here("analysis", "paper", "010-nir-sample-comparison.R"), encoding = "UTF-8"))
```

![Comparison of three dark samples from separate K-cluster groups (model B). Sample 391 from cluster 3 stands out the most with visible differences in both the NIR and Raman spectra, when compared to the other samples](../figures/010-nir-sample-comparison.jpeg){#fig-nir-sample-comparison width=7.87in, height=7.87in}

\linebreak

# Discussion {-}

\linebreak
 \linebreak

The high representation of quartz and quartzite in the Mesolithic - Bronze Age lithic assemblages of Northern Sweden necessitates the development of methods for the characterization of these materials. They represent a primary lithic resource, and thus hold a critical position for understanding resource procurement, trade and exchange, and human mobility within the region (as well as other regions [e.g. @prieto_etal19]). Previous archaeological studies in Northern Sweden have primarily relied on visual characterization, following a system that was devised for managing the large quantities of data generated following the large-scale development of hydroelectric power in the 1940s - 1980s [@biornstad06; @kaeck09: 53-55]. The division of a heterogeneous material into larger simplified colour groups might, however, give a mistaken impression of homogeneity. For example, light grey points with streaks of black mineral running through them might be classified as either light or dark quartzite on different occasions and by different archaeologists. The necessity for adopting such classifications in the past is understandable; a certain level of abstraction is necessary in higher level modelling (Sjölander 2022). Although this work provided plenty of new insights [e.g. @forsberg85; @holm91; @lundberg97], spectroscopic screening provides a more quantitative and empirical approach [@ramacciotti_etal22; @sciuto18].

A benefit of non-destructive screening is the relative ease with which large datasets can be generated. Although there are destructive alternatives that provide a higher resolution view of the material’s chemical characteristic,  non-destructive screening lends itself well to exploratory data analysis and hypothesis generation. The current study demonstrates the potential of new scientific analysis methods for re-assessing and complementing previous research, and thus increasing our understanding of human resource management of locally available lithic material.

A non-destructive approach is particularly useful in regions with complex geology, such as Northern Sweden, where little development has been made in the chemical characterization of quartzite. The results presented here have increased the state of knowledge of the chemical and structural variation that exists under the umbrella of what archaeologists refer to as quartzite within the Västerbotten area, with applications in neighbouring regions and beyond. The material exhibits considerable variation in terms of texture and colour, and the results show that the initial classification, based on the archaeological documentation, may have included mistakes. These mistakes, whilst of little consequence for site interpretation individually, may introduce considerable noise, or “messiness”, when aggregating data for large scale spatio-temporal analyses [@gattiglia15: 2].

The metamorphic nature of quartzite presents a particular challenge for spectroscopic sampling. The material frequently exhibits mineral inclusions that produce a varied surface, and the spot size of the probe will affect the user’s ability to capture a representative area of the artefact. Likewise, the artefact’s shape, type, and preservation significantly influence the availability of surfaces suitable for sampling. Furthermore, dark materials, where a high level of absorbance may saturate measurements or obscure spectral features, are potentially problematic for both NIRS and Raman. Although some of these issues may be overcome with other non-destructive methods, such as Hyperspectral Imaging [@sciuto_etal19], the arguably more affordable techniques presented here produce valuable data.

The spectroscopic features identified in this study relate to the diagenetic/paragenetic environment, and are thus useful for locating potential geological sources in future field surveys. As quartzite is a metamorphic rock, there are two factors affecting its chemical and structural properties; the composition of the materials in the original depositional context, and the grade of metamorphism (in the case of sedimentary orthoquartzite it is mainly the former). The D/G bands indicating graphite in the Raman results imply that the material formed in carbon-rich formations. Alum shales and similar carbon-rich tectonic strata are present within the Scandian mountain chain [@stephens_bergmanweihed20: 482-599], and it would be reasonable to expect carbon structures in quartzite or quartzite-like materials in the region. However, the term “graphitic quartzite” is seldom encountered in contemporary Swedish geological publications. Usage of the term can be found in an early 20th century description of an excursion to Storfjället, Tärna parish, by the Geological Society [@geologiska25]. The occurrence of metamorphic rocks containing graphite, including schist and quartzite, in other regions [e.g. @gaweda_cebulak99; @ukar_cloos16] suggests that further studies of this material would be useful for provenance studies in Sweden.

The metal-OH features in the NIRS data can likewise be considered in terms of metamorphism. If these are related to the Al content of the material and the presence of mica, the band positions may indicate differing metamorphic grades in the material (muscovite - biotite). However, this is not possible to confirm without further statistical analyses, such as multi-blocking [@mishra_etal21], integrating the NIRS and XRF data. Pursuing this angle further could potentially provide useful information for investigating the material’s provenance.

Whilst spectroscopic methods can provide new data for understanding patterns in resource procurement in geologically complex areas, procurement is only part of the picture for resource management strategies during the Mesolithic – Bronze Age. These data can, however, be used to test the different scenarios which existing models represent [e.g. @forsberg85; @holm91; @lundberg97]. For example, a transhumant settlement system in which inland groups, organised along the river valleys, utilize raw materials collected in the mountains, should leave evidence of material flow from the mountains to the forest inland. With enough spectroscopic data from known quartzite artefacts, a predictive model could be built to forecast the type of material expected at other sites. Multiple hypotheses and models could then be tested, and revised, on the basis of further data from new archaeological excavations and geological exploration. The uncertainty of the detailed geology of the region, and the potential noise that glacial disturbance will have introduced, would require that multiple scenarios are tested. A more robust geological resource model could be developed by fusing data from multiple types of spectroscopic instrumentation (e.g. multiblock PCA; Mishra et al., [-@mishra_etal21]).

\linebreak

# Conclusion {-}

\linebreak
 \linebreak

This study provides a case study of a non-destructive and exploratory approach to the characterization of petrographic material where reference material is unavailable. It provides a step forward in the development of methods for the study of quartzite in an archaeological context, and a template for methodologies that can be applied in regions with complex geology and ambiguous distribution of potential geological sources. The results demonstrate clearly that the non-destructive application of all three instrumentations (NIRS, Raman, XRF) is capable of petrographically characterizing some aspect of the quartz and quartzite used in Mesolithic – Bronze Age artefacts in Northern Sweden. Used on their own, however, the individual methods lack the resolution needed to provide a fingerprint detailed enough for classification into distinctive groups. Nevertheless, the methods do individually provide a more in-depth understanding of the composition and structure of the material in comparison to earlier visual studies, as well as showcase the complexity in the petrographic characters of quartz and quartzite materials. The results also demonstrate how the short integration time of these instruments makes it possible to build large datasets that can be statistically explored and mined for information, and ultimately used as the foundation for predictive models for investigating the spatiotemporal aspects of raw material management. This would require the integration of archaeological and geological modelling, including theories of resource procurement and economic strategies, with ice sheet and glacial flow dynamics and detailed Quaternary geological mapping.

\linebreak

# Acknowledgements {-}

We would like to thank Västerbottens museum (www.vbm.se) for access to their collections, and Susanne Sundström in particular for her assistance with preparing the artefacts. We would also like to acknowledge cooperation with the division of Biomass Technology and Chemistry at the Swedish University of Agricultural Sciences.

\linebreak

# Author Contributions {-}
Mattias Sjölander: Conceptualization, Data curation, Formal analysis, Methodology, Project administration, Resources, Software, Visualization, Validation, Writing - original draft, Writing - Review & Editing.

Johan Linderholm: Conceptualization, Methodology, Supervision

Paul Geladi: Formal analysis, Methodology

Philip I. Buckland: Writing – original draft, Writing - Review & Editing.

\linebreak

# Statements and Declarations

**Competing interests**

The authors declare no competing interests.

\linebreak

# Funding source {-}
This work was supported by PhD funding and facilities at the Environmental Archaeology Laboratory at Umeå University.

\linebreak

# Data availability
The code and data used to perform the analysis is openly accessible in an online repository at https://doi.org/10.17605/OSF.IO/F6H98.

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```